---
title: "Functional Data Analysis in Sports Biomechanics"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
knitr
```

```{r, echo=FALSE, include = FALSE}
htmltools::img(src = knitr::image_uri("logo/isbs-logo.png"), 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; padding:10px;',
               width = "150px",
               heigth = "150px")
```

```{r, echo=FALSE, include = FALSE}
htmltools::img(src = knitr::image_uri("logo/fda-logo.png"), 
               alt = 'logo', 
               style = 'position:absolute; top:0; left:0; padding:10px;',
               width = "170px",
               heigth = "170px")
```

# Welcome

This is the web page for the [ISBS 2024](https://www.isbs2024.com) pre-conference workshop [_**"Functional Data Analysis in Sports Biomechanics"**_](https://www.isbs2024.com/wp-content/uploads/2024/03/Pre-workshop_FDA.pdf), delivered by [Prof. Drew Harrison (University of Limerick)](https://www.ul.ie/shprc/professor-drew-harrison) and [Dr. Edward Gunning (University of Pennsylvania)](https://edwardgunning.github.io/).


# 🖥 Computing Pre-requisites

## R and RStudio

You should bring your own laptop with the following software installed:

* **The R Language for Statistical Computing**
  * It can be downloaded from [https://cloud.r-project.org](https://cloud.r-project.org)
  * For further assistance see [this video by RStudio education](https://vimeo.com/203516510)

* **The RStudio Integrated Development Environment (IDE)**
  * It can be downloaded from [https://posit.co/](https://posit.co/)
  * For further assistance see [this video by RStudio education](https://vimeo.com/203516510) (**Note**: The RStudio company has changed to Posit PBC, so there may be some minor differences)


**Note**: If you are unable to install R and RStudio, you can work with a free, lite web version of RStudio called [*posit cloud*](https://posit.cloud/). Watch [this video from Posit PBC](https://www.youtube.com/watch?v=-fzwm4ZhVQQ) to set up an account and get started.

## R Packages

For this workshop, we will primarily use the `fda` (Ramsay, Graves and  Hooker, 2021) and `refund` (Goldsmith et al., 2022) packages.

To install these, you should run the following commands:

```{r, eval=FALSE}
install.packages("fda") # install the fda package
install.packages("refund") # install the refund package
```


---

# 📒 Material
* [**Lecture slides**](slides/template-slides.pptx)

* **Practical Material**
  * [Part 1 -- Data representation and smoothing](practicals/01-smoothing.md)
  * [Part 2 -- Registration](practicals/02-registration.md)
  * Part 3 -- FPCA
  * Part 4 -- Functional regression

---

# 📧 Contact

* Queries about registration for the course should be sent to the ISBS 2024 organisers.

* Queries about the course material should be sent to [edward.gunning@pennmedicine.upenn.edu](mailto:edward.gunning@pennmedicine.upenn.edu) with the subject line *"ISBS pre-conference workshop material"*.

---

# 📚 References

* J. O. Ramsay, Spencer Graves and Giles Hooker (2021). fda: Functional
  Data Analysis. R package version 5.5.1.
  https://CRAN.R-project.org/package=fda
  
* Jeff Goldsmith, Fabian Scheipl, Lei Huang, Julia Wrobel, Chongzhi Di,
  Jonathan Gellar, Jaroslaw Harezlak, Mathew W. McLean, Bruce Swihart,
  Luo Xiao, Ciprian Crainiceanu and Philip T. Reiss (2022). refund:
  Regression with Functional Data. R package version 0.1-26.
  https://CRAN.R-project.org/package=refund


---

# 📖 Further Reading

* Crainiceanu, C. M., Goldsmith, J., Leroux, A., & Cui, E. (2024). Functional Data Analysis with R (1st edition). Chapman and Hall/CRC (book website: [https://functionaldataanalysis.org](https://functionaldataanalysis.org))

* Ramsay, J. O., & Silverman, B. W. (2005). Functional Data Analysis (2nd ed.). Springer-Verlag. https://doi.org/10.1007/b98888

* Ramsay, J. O., Hooker, G., & Graves, S. (2009). Functional Data Analysis with R and MATLAB. Springer-Verlag. https://doi.org/10.1007/978-0-387-98185-7


